Order statistics and numerical optimization: the case of a single observation

### 2020-01-26 inst/order.Rmd
more work on the limiting process
more work

---
title: "Order statistics and numerical optimization"
output: html_document
---

@@

\[
\lim_{p\longrightarrow 0}
{\mathcal L}(1-9p,p,p,p,p,p,p,p,p,p,p)=\frac{1}{8!}
\]

but

\[
\lim_{p\longrightarrow 0}
{\mathcal L}(\alpha,\alpha p,\alpha p^2,\ldots,\alpha p^9)=1
\]

where $\alpha=(1-p^{10})/(1-p)$ is a normalizing constant.  Thus the
likelihood _at_ the evaluate is not defined, even as a limit (because
it depends on the direction of approach).

@@

```{r}
dotchart(evaluate2,pch=16)
```

So it is much better than the coldstart above in the sense that there
is more probability on $p_1$ and less everywhere else.  We can compare
the hotstart above to the coldstart:

@@

This is negative: the so-called "sensible start point" has produced an
inferior solution.  This is because `sss` has $p_2=p_3=\ldots =
p_{10}$, so predicts a draw between players $2,3,\ldots 10$, contrary
to observation.

A better way is to mimic the more appropriate of the two limiting processes above:

@@

Observe that the likelihood at this point is marginally better than
the `equalp()` start point:

@@

but the dotchart again shows a quite a lot of probability on
$p_2,\ldots p_{10}$.

>>>>

---
title: "Order statistics and numerical optimization: the case of a single observation"
output: html_document
---

@@

\[
\lim_{x\longrightarrow 0}
{\mathcal L}(1-9x,x,x,x,x,x,x,x,x,x,x)=\frac{1}{8!};\qquad
\lim_{x\longrightarrow 0}
{\mathcal L}(\alpha,\alpha x,\alpha x^2,\ldots,\alpha x^9)=1
\]

where $\alpha=(1-x^{10})/(1-x)$ is a normalizing constant.  Thus the
likelihood _at_ the evaluate is not defined, even as a limit (because
it depends on the direction of approach).  However, it is reasonable
to demand consistency in the limiting process.  Thus, writing
$p_i(x),1\leq i\leq 10$, we require

\[
\lim_{x\longrightarrow 0}\frac{p_i(x)}{p_{i+1}(x)}=0,\qquad i=1,\ldots,9
\]

and if this is respected, then it is reasonable to say that the
likelihood at the evaluate is 1, and the support 0.

## Package idiom for likelihood functions

@@

```{r}
dotchart(evaluate2,pch=16)
par(pty='s')
plot(evaluate1,evaluate2,asp=1,log='xy')
abline(0,1)
```

So it is almost identical to the coldstart above in the sense that the
two evaluates are very close.  We can compare the hotstart to the
coldstart in terms of support:

@@

This is almost exactly zero, but negative: the so-called "sensible
start point" has produced an inferior solution.  This is because `sss`
has $p_2=p_3=\ldots = p_{10}$, so predicts a draw between players
$2,3,\ldots 10$, contrary to observation.

Possibly, a better way is to mimic the more appropriate of the two
limiting processes above:

@@

Observe that the likelihood at this point is only very slightly better
than the `equalp()` start point:

@@
